---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message=FALSE, warning=FALSE}
devtools::load_all(".")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


This package is used to analyze Serie A soccer (Calcio) data. It creates an accessible R data-frame with information about match results, as well as team stats, Elo ratings, and overall standings. This data-frame is used to generate visualizations on a Shiny App: https://datavisr.shinyapps.io/calcior/




#### Source Data
The data is sourced from https://github.com/openfootball which contains the results of all Serie A match since the 2013/14 season. The data is extracted using Ruby with the sportdb gem. Running this will create a local SQLite database `sport.db` that we can use to read into R.

```{r, echo=FALSE}
    source_data <- dao$new()
    source_data$get_data()
    filter_na_cols <- function(df) df[,purrr::map_lgl(df, ~!all(is.na(.x)))]
    filter_at_cols <- function(df) df %>% select(-one_of("created_at", "updated_at"))
    source_data$tables %>% 
        purrr::map(filter_na_cols) %>% 
        purrr::map(filter_at_cols) %>% 
        glimpse()

```

#### Processed Data

The source data is transformed from a set of relational tables to a single data-frame `serie_a` which contains list columns of data-frame to maintain the relationship of teams and matches to match_days (rounds) and season. Summary data and Elo ratings are also calculated (details below).


```{r}
serie_a
```
##### `season`:
Serie A seasons starting from 2013/14 to 2016/17

##### `match_days_complete`:
The number of matches completed so far for each season.

##### `teams`:
The teams included for each season in Serie A. They change each season as the bottom 3 teams are sent down to Serie B and the top 3 teams from Serie B are promoted.

```{r}
serie_a %>% select(season, teams) %>% tidyr::unnest(teams) %>% glimpse()
```



##### `results`: 
For every `season`, `match_day` and team (`p_team`, for primary team) it shows their score (`p_score`), their opponents score (`o_score`), if they were home (`p_home`) and how many `points` the `p_team` earned from the result. 
```{r}
serie_a %>% select(season, results) %>% tidyr::unnest(results) %>% tidyr::unnest(data) %>% glimpse()
```


##### `ratings`:
For every `season`, `match_day` and team (`p_team`) it shows the teams Elo rating `r`. 

The Elo calculations are mostly based on this site: http://www.eloratings.net/system.html. With `k` = 20 and a season reverting factor of 0.25. 
```{r}
serie_a %>% select(season, ratings) %>% tidyr::unnest(ratings) %>% tidyr::unnest(data) %>% glimpse()
```


##### `standings`:
For every `season`,` match_day` and team (`p_team`) it shows the teams cumulative `points`, `goals_for`, `goals_against` and `goal_diff`, along with their `position` in comparison to the other teams.
```{r}
serie_a %>% select(season, standings) %>% tidyr::unnest(standings) %>% tidyr::unnest(data) %>% glimpse()
```
